---
title: "esm201_ass2"
author: "Jeremy Knox"
date: "3/9/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, 
                      cache=FALSE, 
                      fig.width=5, 
                      fig.height=5,
                      fig.align='center')
indent1 = '    '      
indent2 = paste(rep(indent1, 2), collapse='')

library(tidyverse)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
library(stargazer)
library(latex2exp)
library(finalfit)

fert = read_csv("fertilizer.csv")
grain = read_csv("grain.csv")

```


## 1. Patterns in yield over time  

# a. Plot of yield vs. time by grain 
```{r error=FALSE, echo=FALSE}

grains_plot = ggplot(data = grain, aes(x=year, y=yield)) +
  geom_point() +
  facet_wrap(~commodity, scales = "free") +
  theme_classic() +
  xlab(TeX('Year')) +
  ylab(TeX('Grain Yield (bushels $acre^{-1}$) ')) 
grains_plot

```
**Figure 1 | Four plots showing grain yeild over time by commodity.** Grain yeild is measured in bushels per acre. Barley and Corn look to have a linear low plateau for yield over time with  positive slope starting in the 1940s. Oats and Sorghum look to have a linear piecewise relationship for yiled over time with an increasing rate starting in the 1950s and 1960s, respectivley.

Barley and Corn yield over time, seen in *Figure 1*, could be described by a linear low plateau model. For example, let $x =$ year and $y =$ Grain Yield (bushels   acre$^{-1}$). Then for Barley we have:
$$ 
\begin{cases} 
      y_0 \approx 25 & x\leq 1950 \\
      \ y \ \approx y_0 + \frac{4}{5}x & x > 1950 
   \end{cases}
$$
